Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-02-21 paper/ms.tex
a few words about the 'wrong' example [ci skip]

% Projects:
\newcommand{\project}[1]{\textsf{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\celerite}{\project{celerite}}

@@

It's worth noting that the product of two terms of the form found inside the
sum in \eq{celerite-kernel} can also be re-written as a sum with updated
parameters
\begin{eqnarray}
k_j(\tau) \, k_k(\tau) =
    e^{-\tilde{c}\,\tau}\,[
        a_+\,\cos(d_+\,\tau) + b_+\,\sin(d_+\,\tau) +
        a_-\,\cos(d_-\,\tau) + b_-\,\sin(d_-\,\tau)
    ]
\end{eqnarray}
where
\begin{eqnarray}
    \tilde{a}_{\pm} &=& \frac{1}{2}\,(a_j\,a_k \pm b_j\,b_k) \\
    \tilde{b}_{\pm} &=& \frac{1}{2}\,(b_j\,a_k \mp a_j\,b_k) \\
    \tilde{c} &=& c_j + c_k \\
    \tilde{d}_{\pm} &=& d_j \mp d_k \quad.
\end{eqnarray}

@@

These identities yield a kernel of the form
\begin{equation}
k(\tau) = S_0\,\omega_0\,Q\,e^{-\frac{\omega_0\,\tau}{2Q}}\,
\begin{cases}
    \cosh{(\eta\,\omega_0\,\tau)} +
        \frac{1}{2\,\eta\,Q}\,\sinh{(\eta\,\omega_0\,\tau)}, & 0 < Q < 1/2\\
    2\,(1+\omega_0\,\tau), & Q = 1/2\\
    \cos{(\eta\,\omega_0\,\tau)} +
        \frac{1}{2\,\eta\,Q} \sin{(\eta\,\omega_0\,\tau)},& 1/2 < Q\\
\end{cases}
\end{equation}
where $\eta = \vert 1-(4\,Q^2)^{-1}\vert^{1/2}$.
It is interesting to note that, because of the damping, the characteristic
oscillation frequency in this model $d_j$, for any finite quality factor $Q > 1/2$,
is not equal to the frequency of the undamped oscillator $\omega_0$.

@@

>>>>

% Projects:
\newcommand{\project}[1]{\textsf{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\celerite}{\project{celerite}}
\newcommand{\emcee}{\project{emcee}}

@@

It's worth noting that the product of two terms of the form found inside the
sum in \eq{celerite-kernel} can also be re-written as a sum with updated
parameters
\begin{eqnarray}\eqlabel{product-rule}
k_j(\tau) \, k_k(\tau) =
    e^{-\tilde{c}\,\tau}\,[
        a_+\,\cos(d_+\,\tau) + b_+\,\sin(d_+\,\tau) +
        a_-\,\cos(d_-\,\tau) + b_-\,\sin(d_-\,\tau)
    ]
\end{eqnarray}
where
\begin{eqnarray}
    \tilde{a}_{\pm} &=& \frac{1}{2}\,(a_j\,a_k \pm b_j\,b_k) \\
    \tilde{b}_{\pm} &=& \frac{1}{2}\,(b_j\,a_k \mp a_j\,b_k) \\
    \tilde{c} &=& c_j + c_k \\
    \tilde{d}_{\pm} &=& d_j \mp d_k \quad.
\end{eqnarray}

@@

These identities yield a kernel of the form
\begin{equation}\eqlabel{sho-kernel}
k(\tau) = S_0\,\omega_0\,Q\,e^{-\frac{\omega_0\,\tau}{2Q}}\,
\begin{cases}
    \cosh{(\eta\,\omega_0\,\tau)} +
        \frac{1}{2\,\eta\,Q}\,\sinh{(\eta\,\omega_0\,\tau)}, & 0 < Q < 1/2\\
    2\,(1+\omega_0\,\tau), & Q = 1/2\\
    \cos{(\eta\,\omega_0\,\tau)} +
        \frac{1}{2\,\eta\,Q} \sin{(\eta\,\omega_0\,\tau)},& 1/2 < Q\\
\end{cases}
\end{equation}
where $\eta = \vert 1-(4\,Q^2)^{-1}\vert^{1/2}$.
It is interesting to note that, because of the damping, the characteristic
oscillation frequency in this model $d_j$, for any finite quality factor $Q > 1/2$,
is not equal to the frequency of the undamped oscillator $\omega_0$.

@@

For this example, the data are simulated from a quasiperiodic GP with the
kernel
\begin{eqnarray}\eqlabel{sim-wrong-true}
k_\mathrm{true} (\tau) = \alpha\,
    \exp\left(-\frac{\tau^2}{2\,\lambda^2}\right)\,
    \cos\left(\frac{2\,\pi\,\tau}{P_\mathrm{true}}\right)
\end{eqnarray}
where $P_\mathrm{true}$ is the fundamental period of the process.
This autocorrelation structure yields the power spectrum
\begin{eqnarray}
S_\mathrm{true} (\omega) = \frac{\lambda\,\alpha}{2}\,\left[
    \exp\left(-\frac{\lambda^2}{2}\,\left(\omega-
        \frac{2\,\pi}{P_\mathrm{true}}\right)^2\right) +
    \exp\left(-\frac{\lambda^2}{2}\,\left(\omega+
        \frac{2\,\pi}{P_\mathrm{true}}\right)^2\right)
\right]
\end{eqnarray}
that for large $\omega$ falls off exponentially.
When compared to \eq{celerite-psd}~--~that for large $\omega$ goes as
$\omega^{-4}$, at most~--~it is clear that a \celerite\ model can never
perfectly reproduce the structure of this process.
That being said, we will demonstrate that rigorous inferences can be made
about $P_\mathrm{true}$ even with an effective model.
The left panel of \Figure{simulated-wrong} shows the simulated dataset.
We then fit this simulated data using the product of two SHO terms
(\eqalt{sho-kernel}) where one of the terms has $S_0 = 1$ and $Q =
1/\sqrt{2}$ and the other has $\omega_0 = 2\,\pi/P$.
We note that using \eq{product-rule}, the product of two \celerite\ terms can
also be expressed as a \celerite\ term.
We apply log-uniform priors on all the parameters and use \emcee\
\citep{Foreman-Mackey:2013} to sample the posterior probability for all of the
parameters.
The inferred distribution for the parameter $P$ is shown in the right panel of
\Figure{simulated-wrong} and compared to the true period $P_\mathrm{true}$ and
the inferences made using the correct model (\eqalt{sim-wrong-true}).
The inference made using this effective \celerite\ model are indistinguishable
from the inferences made using the correct model but substantially less
computation time is required for the \celerite\ inference.
